---
title: "Energy Load Stylized Facts"
author: "JW"
date: "`r Sys.Date()`"
output: html_document

---
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
```

Select the years to analyse!
```{r}
energy_load <- read.csv("../data/load_15-24.csv")
energy_load$date <- as.POSIXct(energy_load$date, tz = "UTC")
```


# Energy Load
Summary statistics
```{r}
summary(energy_load)

paste("SD for load: ", sapply(energy_load, sd, na.rm = TRUE)[3], " MWh")
```

Overview of the whole time series
```{r, eval=FALSE}
energy_load |>
    ggplot(aes(date, load)) +
    geom_line() +
    labs(
        x = "Time",
        y = "Load in MWh",
        title = "Loads over the years 2015 - 2024",
    )
```

The average load per hour for the whole time series
```{r}
energy_load |>
    group_by(hour_int) |>
    summarise(mean_load = mean(load)) |>
    ggplot(aes(x = hour_int, y = mean_load)) +
    geom_point() +
    geom_line() +
    labs(
        x = "Hour of the day",
        y = "Mean Load",
        title = "Average Load over the years 2015-2024 per day",
    )
```

The average load per hour grouped by year
```{r}
# Add a new column for the year


# Group by hour and year, then calculate the mean load
avg_load_by_hour_year <- energy_load |>
    mutate(year = year(date)) |>
    group_by(hour_int, year) |>
    summarise(mean_load = mean(load))

# Plot
avg_load_by_hour_year |>
    ggplot(aes(x = hour_int, y = mean_load, colour = factor(year))) +
    geom_point() +
    geom_line() +
    labs(
        x = "Hour of the day",
        y = "Mean Load",
        title = "Average Load over the years 2015-2024 per day",
    )
```

Boxplot of the load per day for the weeks from 2022 - 2024
```{r}
energy_load |>
    ggplot(aes(x = as.factor(weekday_int), y = load)) +
    geom_boxplot() +
    labs(x = "Weekday", y = "Load in MWh", title = "Load per weekday from 2015 - 2024")
```

Boxplot of the load per month for periode from 2022 - 2024

```{r}
energy_load |>
    ggplot(aes(x = as.factor(month_int), y = load)) +
    geom_boxplot() +
    labs(x = "Month", y = "Load in MWh", title = "Load per month from 2015 - 2024")
```

Load over the whole time period by months and grouped by working day
```{r}
month <- 2
energy_load |>
    filter(month_int == month) |>
    ggplot(aes(hour_int, load, color = working_day)) +
    geom_point() +
    labs(
        x = "Hour of the day",
        y = "Load in MWh",
        title = "Load per hour for the period 2015 - 2024",
        subtitle = paste("grouped by working day and filtered by month ", month)
    )
```

Overview of a time series of loads per Month for a given year and month
```{r}
# --- Get the time series of loads for a given year and month ---
year <- 2024
month <- 1
# Create the figure
energy_load |>
    filter(year(date) == year & month_int == month) |>
    mutate(week_day_hour = weekday_int + hour_int / 24, week_int = week(date)) |>
    ggplot(aes(x = week_day_hour, y = load, colour = factor(week_int))) +
    geom_point(size = 0.5) +
    geom_line() +
    scale_x_continuous(breaks = 1:7, labels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")) +
    labs(
        x = "Day of the week",
        y = "Load in MWh",
        colour = "Week of the month",
        title = paste("Load per hour for the year ", year, " and the month ", month)
    )
```

A histogramm of the data for the whole time series
```{r}
ggplot(energy_load, aes(load)) +
    geom_histogram(binwidth = 500) +
    labs(
        x = "Load in MWh",
        title = "Histogramm of all data points from 2015 - 2024",
    )
```

A histogramm of the data grouped by working days and weekends

```{r}
ggplot(energy_load, aes(load)) +
    geom_histogram(binwidth = 500) +
    labs(
        title = "Histogramm of Loads for 2015 - 2024 for Working Days",
        x = "Load in MWh"
    ) +
    facet_wrap(~working_day)
```

# Energy Load Peaks

Getting the peaks
```{r}
peaks <- energy_load |>
    group_by(as.Date(date)) |>
    slice(which.max(load))
```

Overview of peaks over time
```{r}
peaks |>
    ggplot(aes(date, load)) +
    geom_line() +
    labs(
        x = "Time",
        y = "Load in MWh",
        title = "Load peaks over the years 2015 - 2024",
    )
```

Histogramm of the peaks
```{r}
peaks |>
    ggplot(aes(x = load)) +
    geom_histogram(binwidth = 250) +
    labs(
        title = "Histogramm of load peaks from 2015 to 2024",
        x = "Load in MWh"
    )
```

Histogramm of the peaks grouped by working days
```{r}
peaks |>
    ggplot(aes(load)) +
    geom_histogram(binwidth = 250) +
    labs(
        title = "Histogramm of Loads for 2015 - 2024 for Working Days",
        x = "Load in MWh"
    ) +
    facet_wrap(~working_day)
```

Histogramm of the peaks based on hours of the day
```{r}
peaks |>
    ggplot(aes(hour_int)) +
    geom_histogram(binwidth = 1) +
    labs(
        title = "Histogramm of Peak Loads for 2015 - 2024 per hour of the day",
        x = "Hour of the day"
    )
```

Scatterplot of the load peaks per hour grouped by working days 
```{r}
peaks |>
    ggplot(aes(x = hour_int, y = load, color = working_day)) +
    geom_point() +
    scale_x_continuous(breaks = seq(min(peaks$hour_int), max(peaks$hour_int), by = 1)) +
    labs(
        title = "Scatterplot of Load peaks from 2015 - 2024",
        x = "Hour of the day",
        y = "Load in MWh"
    )
```

Density of the peaks
```{r}
ggplot(peaks, aes(x = load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks 2015 - 2024",
        x = "Load in MWh"
    )
```

Density of the peaks grouped by working days
```{r}
ggplot(peaks, aes(x = load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks 2015 - 2024 grouped by working days",
        x = "Load in MWh"
    ) +
    facet_wrap(~working_day)
```

Boxplot of the peaks per day for the weeks from 2022 - 2024
```{r}
peaks |>
    ggplot(aes(x = as.factor(weekday_int), y = load)) +
    geom_boxplot() +
    labs(x = "Weekday", y = "Load in MWh", title = "Daily peak per weekday from 2015 - 2024")
```

Boxplot of the peaks per month for periode from 2022 - 2024

```{r}
peaks |>
    ggplot(aes(x = as.factor(month_int), y = load)) +
    geom_boxplot() +
    labs(x = "Month", y = "Load in MWh", title = "Peak load per month from 2015 - 2024")
```

## Bivariate Density of Load peaks
```{r, message=FALSE}
library(MASS)
library(plotly)
```

Density of **energy load** on hourly resolution
```{r}
den3d <- kde2d(energy_load$load, energy_load$hour_int)
plot_ly(x = den3d$x, y = den3d$y, z = den3d$z) |> add_surface()
```

Density of **load peaks** per day for working days
```{r}
peaks_working <- peaks |>
    filter(working_day == TRUE)

peaks_working |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks for the day grouped by hour for working days",
        x = "Load in MWh"
    ) +
    facet_wrap(~hour_int)
```

Depicting the density of hour 11
```{r}
peaks_working |>
    filter(hour_int == 11) |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks on Working days at hour 11",
        x = "Load in MWh"
    )
```

Depicting the density of hour 12
```{r}
peaks_working |>
    filter(hour_int == 12) |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks on Working days at hour 12",
        x = "Load in MWh"
    )
```

Depicting the density of hour 13
```{r}
peaks_working |>
    filter(hour_int == 13) |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks on Working days at hour 13",
        x = "Load in MWh"
    )
```

Density of **load peaks** per day for weekends
```{r}
peaks_weekends <- peaks |>
    filter(working_day == FALSE)

peaks_weekends |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks for the day grouped by hour for weekends",
        x = "Load in MWh"
    ) +
    facet_wrap(~hour_int)
```

Depicting the density of hour 18
```{r}
peaks_weekends |>
    filter(hour_int == 18) |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks for the weekends at hour 18",
        x = "Load in MWh"
    )
```

Depicting the density of hour 19
```{r}
peaks_weekends |>
    filter(hour_int == 19) |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks for the weekends at hour 19",
        x = "Load in MWh"
    )
```

Depicting the density of hour 20
```{r}
peaks_weekends |>
    filter(hour_int == 20) |>
    ggplot(aes(load)) +
    geom_density() +
    labs(
        title = "Density of the load peaks for the weekends at hour 20",
        x = "Load in MWh"
    )
```
