---
title: "Shaake Shuffle"
author: "JW"
date: "`r Sys.Date()`"
output: html_document
---


## Load the data and packages
```{r message=FALSE}
library(tidyverse)
library(scoringRules)
library(ggplot2)
library(forecast)
library(dynlm)
library(plotly)

# For the holidays
holidays <- read.csv("../data/holidays_DE_15-24.csv") |>
    mutate_at("Date", as.Date)

# --- Load the energy data ----
energy_load <- read.csv("../data/load_15-24.csv") |>
    mutate_at(c("hour_int", "weekday_int", "month_int"), as.factor) |>
    mutate(is_holiday = if_else(as.Date(date) %in% holidays$Date, 1, 0))

energy_load$date <- as.POSIXct(energy_load$date, tz = "UTC")

# --- Getting the peaks ---
peaks <- energy_load |>
    group_by(as.Date(date)) |>
    slice(which.max(load))
```

We consider the year 2015 for training for 365 days and then predicting onwards...

## General Model Approach A - Univariate
### Model 1: AR1
```{r}
predict_ar1 <- function(data, d) {
    # INPUT:
    # data
    # d ...  day

    # length of the training period is 365 by default
    # Generate lag 1
    data$load_lag_1 <- lag(data$load, 1)
    data <- na.omit(data)
    # Train/Test Split

    train <- data[d:(364 + d), ]
    test <- data[(365 + d), ]

    # ------- TRAINING ----------

    # AR1
    model <- lm(load ~ load_lag_1 + weekday_int + is_holiday, data = train)

    # --------- TESTING ------------
    yhat_test <- as.numeric(predict(model, newdata = test))

    test$y_hat <- yhat_test
    return(test)
}
```

```{r}
# Run the predictons for some days
predictions <- data.frame(matrix(ncol = 9, nrow = 0))
for (i in 1:150) {
    value <- predict_ar1(peaks, i)
    predictions <- rbind(predictions, value)
}
```

```{r}
# Reshape the dataframe to long format
predictions_long <- pivot_longer(predictions, cols = c(load, y_hat), names_to = "type", values_to = "value")

# Plot
fig <- ggplot(predictions_long, aes(x = date, y = value, color = type)) +
    geom_line() + # Draw lines
    labs(x = "Date", y = "Value", title = "Load and Predicted Load Over Time - AR1") +
    theme_minimal() # Use a minimal theme for aesthetics

ggplotly(fig)
```

```{r}
resids <- predictions$load - predictions$y_hat

hist(resids, main = "Histogram of Residuals", xlab = "Residuals - AR1")
```

### Model 2: ARIMA(0,1,2)
```{r}
predict_arima <- function(data, d) {
    # INPUT:
    # data
    # d ...  day

    # length of the training period is 365 by default

    # Train/Test Split
    train <- data[d:(364 + d), ]
    test <- data[(365 + d), ]
    weekday_dummies <- model.matrix(~ factor(weekday_int), data = data)[, -1]

    # ------- TRAINING ----------
    # Set up the X matrix for training
    x_train <- as.matrix(cbind(weekday_dummies[d:(364 + d), ], train$is_holiday))
    # Set column names, assuming the first columns are from weekday_dummies_test and the last is is_holiday
    colnames(x_train) <- c(paste("weekday", 2:7, sep = "_"), "is_holiday")

    # ARIMA
    model <- arima(train$load, c(0, 1, 2), xreg = x_train)

    # --------- TESTING ------------
    x_test <- as.matrix(c(weekday_dummies[(365 + d), ], test$is_holiday))
    # Set column names, assuming the first columns are from weekday_dummies_test and the last is is_holiday
    rownames(x_test) <- c(paste("weekday", 2:7, sep = "_"), "is_holiday")
    yhat_test <- as.numeric(predict(model, n.ahead = 1, newxreg = t(x_test))$pred)

    test$y_hat <- yhat_test
    return(test)
}
```

```{r}
# Run the predictons for some days
predictions <- data.frame(matrix(ncol = 9, nrow = 0))
for (i in 1:150) {
    value <- predict_arima(peaks, i)
    predictions <- rbind(predictions, value)
}
```

```{r}
# Reshape the dataframe to long format
predictions_long <- pivot_longer(predictions, cols = c(load, y_hat), names_to = "type", values_to = "value")

# Plot
fig <- ggplot(predictions_long, aes(x = date, y = value, color = type)) +
    geom_line() + # Draw lines
    labs(x = "Date", y = "Value", title = "Load and Predicted Load Over Time - ARIMA(0,1,2)") +
    theme_minimal() # Use a minimal theme for aesthetics

ggplotly(fig)
```

```{r}
resids <- predictions$load - predictions$y_hat

hist(resids, main = "Histogram of Residuals", xlab = "Residuals - ARIMA(0,1,2)")
```

### Model 3: NeuralProphet
*Already implemented and working well...*


## General Model Approach B - Multivariate
### Model 1: AR24

```{r}
predict_ar24 <- function(data, d) {
    # INPUT:
    # data
    # d ...  day

    # length of the training period is 365 by default

    # --------- Data Preprocessing -----------
    # Create lagged variables for the model
    data <- data |>
        mutate(
            Y_1_h = lag(load, 1),
            Y_2_h = lag(load, 2),
            Y_3_h = lag(load, 3),
            Y_4_h = lag(load, 4),
            Y_5_h = lag(load, 5),
            Y_6_h = lag(load, 6),
            Y_7_h = lag(load, 7),
            Y_8_h = lag(load, 8),
            Y_9_h = lag(load, 9),
            Y_10_h = lag(load, 10),
            Y_11_h = lag(load, 11),
            Y_12_h = lag(load, 12),
            Y_13_h = lag(load, 13),
            Y_14_h = lag(load, 14),
            Y_15_h = lag(load, 15),
            Y_16_h = lag(load, 16),
            Y_17_h = lag(load, 17),
            Y_18_h = lag(load, 18),
            Y_19_h = lag(load, 19),
            Y_20_h = lag(load, 20),
            Y_21_h = lag(load, 21),
            Y_22_h = lag(load, 22),
            Y_23_h = lag(load, 23),
            Y_24_h = lag(load, 24)
        )

    # Create weekday dummy variables
    weekday_dummies <- model.matrix(~ factor(weekday_int), data = data)[, -1]

    for (i in 2:7) {
        data[[paste0("DoW_", i)]] <- weekday_dummies[, i - 1]
    }

    # Remove rows with NA values created by lagging
    data <- na.omit(data)

    # ------- TRAINING ----------
    # Formula for the regression
    formula <- load ~ Y_1_h + Y_2_h + Y_3_h + Y_4_h + Y_5_h + Y_6_h +
        Y_7_h + Y_8_h + Y_9_h + Y_10_h + Y_11_h + Y_12_h +
        Y_13_h + Y_14_h + Y_15_h + Y_16_h + Y_17_h + Y_18_h +
        Y_19_h + Y_20_h + Y_21_h + Y_22_h + Y_23_h + Y_24_h +
        DoW_2 + DoW_3 + DoW_4 + DoW_5 + DoW_6 + DoW_7 + is_holiday

    # Fit the regression model
    # Use One year for training
    train <- data[(24 * d - 23):(((364 + d) * 24)), ]
    model <- lm(formula, data = train)

    # --------- TESTING ------------
    # getting the next hour of data
    test <- data[(24 * d + 365 * 24 - 23):(24 * d + 365 * 24 - 23), ]

    predictions <- numeric(24)
    for (i in 1:24) {
        predictions[i] <- as.numeric(predict(model, newdata = test))

        # Shift Y_1_h to Y_24_h one position to the right and insert the new prediction at Y_1_h
        test <- test %>%
            mutate(
                Y_24_h = Y_23_h,
                Y_23_h = Y_22_h,
                Y_22_h = Y_21_h,
                Y_21_h = Y_20_h,
                Y_20_h = Y_19_h,
                Y_19_h = Y_18_h,
                Y_18_h = Y_17_h,
                Y_17_h = Y_16_h,
                Y_16_h = Y_15_h,
                Y_15_h = Y_14_h,
                Y_14_h = Y_13_h,
                Y_13_h = Y_12_h,
                Y_12_h = Y_11_h,
                Y_11_h = Y_10_h,
                Y_10_h = Y_9_h,
                Y_9_h = Y_8_h,
                Y_8_h = Y_7_h,
                Y_7_h = Y_6_h,
                Y_6_h = Y_5_h,
                Y_5_h = Y_4_h,
                Y_4_h = Y_3_h,
                Y_3_h = Y_2_h,
                Y_2_h = Y_1_h,
                Y_1_h = predictions[i]
            )
    }
    data_test <- data[(24 * d + 365 * 24 - 23):(24 * d + 365 * 24), ]
    data_test$y_hat <- predictions
    return(data_test)
}
```

```{r}
# Run the predictons for some days
predictions <- data.frame(matrix(ncol = 38, nrow = 0))
for (i in 1:21) {
    value <- predict_ar24(energy_load, i)
    predictions <- rbind(predictions, value)
}
```

```{r}
# Reshape the dataframe to long format
predictions_long <- pivot_longer(predictions, cols = c(load, y_hat), names_to = "type", values_to = "value")

# Plot
fig <- ggplot(predictions_long, aes(x = date, y = value, color = type)) +
    geom_line() + # Draw lines
    labs(x = "Date", y = "Value", title = "Load and Predicted Load Over Time - ARIMA(0,1,2)") +
    theme_minimal() # Use a minimal theme for aesthetics
ggplotly(fig)
```

```{r}
resids <- predictions$load - predictions$y_hat

hist(resids, main = "Histogram of Residuals", xlab = "Residuals - AR24")
```

### Model 2: ARIMA-Fourier

```{r eval=FALSE}
train <- msts(data[1:(365 * 24), ]$load, seasonal.period = c(24, 7 * 24))

# ------- Fourier Model -----
z <- fourier(train, K = c(5, 5))
holidays <- data[1:(365 * 24), ]$is_holiday
x_train <- cbind(z, holidays)

# Time duration to fit: 102 sec.
fit <- auto.arima(train, seasonal = FALSE, xreg = cbind(z, holidays))

zf <- fourier(train, K = c(5, 5), h = 24)
holidaysf <- data[((365 * 24) + 1):((365 * 24) + 24), ]$is_holiday

predictions <- as.numeric(predict(fit, newxreg = cbind(zf, holidaysf))$pred)

# Assuming 'date' is the vector of dates corresponding to your test data
# Create a new data frame for plotting
plot_data <- data.frame(date = data[((365 * 24) + 1):((365 * 24) + 24), ]$date, Actual = data[((365 * 24) + 1):((365 * 24) + 24), ]$load, Predicted = predictions)

# Melt the data frame for easier plotting with ggplot
plot_data_long <- reshape2::melt(plot_data, id.vars = "date")

# Plot
ggplot(plot_data_long, aes(x = date, y = value, color = variable)) +
    geom_line() +
    labs(title = "Actual vs Predicted Load", y = "Load", x = "Date") +
    theme_minimal() +
    scale_color_manual(values = c("Actual" = "red", "Predicted" = "blue"))
```

### Model 3: ARIMA
```{r}
predict_arima <- function(data, d) {
    # INPUT:
    # data
    # d ...  day

    # length of the training period is 365 days by default

    # Train/Test Split
    # Take 365 days from day d on - Remeber we have 24 hours = 1 day
    train <- data[(24 * d - 23):(((364 + d) * 24)), ]
    # Get the next 24 hours after the testing period
    test <- data[(24 * d + 365 * 24 - 23):(24 * d + 365 * 24), ]

    weekday_dummies <- model.matrix(~ factor(weekday_int), data = data)[, -1]

    # ------- TRAINING ----------
    # Set up the X matrix for training
    x_train <- as.matrix(cbind(weekday_dummies[(24 * d - 23):(((364 + d) * 24)), ], train$is_holiday))
    # Set column names, assuming the first columns are from weekday_dummies_test and the last is is_holiday
    colnames(x_train) <- c(paste("weekday", 2:7, sep = "_"), "is_holiday")

    # ARIMA - CHANGE MODEL here
    model <- arima(train$load, c(3, 1, 3), xreg = x_train)


    # --------- TESTING ------------
    x_test <- as.matrix(cbind(weekday_dummies[(24 * d + 365 * 24 - 23):(24 * d + 365 * 24), ], test$is_holiday))
    # Set column names, assuming the first columns are from weekday_dummies_test and the last is is_holiday
    colnames(x_test) <- c(paste("weekday", 2:7, sep = "_"), "is_holiday")

    # Predict new values with MODEL
    yhat_test <- as.numeric(predict(model, n.ahead = 24, newxreg = x_test)$pred)

    test$y_hat <- yhat_test
    return(test)
}
```

```{r}
# Run the predictons for some days
predictions <- data.frame(matrix(ncol = 9, nrow = 0))
for (i in 1:14) {
    print(i)
    value <- predict_arima(energy_load, i)
    predictions <- rbind(predictions, value)
}
```
```{r}
# Reshape the dataframe to long format
predictions_long <- pivot_longer(predictions, cols = c(load, y_hat), names_to = "type", values_to = "value")

# Plot
fig <- ggplot(predictions_long, aes(x = date, y = value, color = type)) +
    geom_line() + # Draw lines
    labs(x = "Date", y = "Value", title = "Load and Predicted Load Over Time - ARIMA(3,1,3)") +
    theme_minimal() # Use a minimal theme for aesthetics
fig

resids <- predictions$load - predictions$y_hat

hist(resids, main = "Histogram of Residuals", xlab = "Residuals ARIMA")
```

### Model 4: Expert Model
```{r}
predict_ar24 <- function(data, d) {
    # INPUT:
    # data
    # d ...  day

    # length of the training period is 365 by default

    # --------- Data Preprocessing -----------
    # Create lagged variables for the model
    data <- data |>
        mutate(
            Y_d_1_h = lag(load, 1 * 24), # lag by 24 hours
            Y_d_2_h = lag(load, 2 * 24), # lag by 48 hours
            Y_d_7_h = lag(load, 7 * 24), # lag by 168 hours (1 week)
            Y_d_1_min = sapply(1:n(), function(i) ifelse(i > 24, min(load[(i - 24):(i - 1)]), NA)),
            Y_d_1_max = sapply(1:n(), function(i) ifelse(i > 24, max(load[(i - 24):(i - 1)]), NA)),
            Y_d_1_24 = lag(load, 1)
        )

    # Create weekday dummy variables
    weekday_dummies <- model.matrix(~ factor(weekday_int), data = data)[, -1]

    for (i in 2:7) {
        data[[paste0("DoW_", i)]] <- weekday_dummies[, i - 1]
    }

    # Remove rows with NA values created by lagging
    data <- na.omit(data)

    # ------- TRAINING ----------
    # Formula for the regression
    formula <- load ~ Y_d_1_h + Y_d_2_h + Y_d_7_h + Y_d_1_min + Y_d_1_max + Y_d_1_24 +
        DoW_2 + DoW_3 + DoW_4 + DoW_5 + DoW_6 + DoW_7

    # Fit the regression model
    # Use One year for training
    train <- data[(24 * d - 23):(((364 + d) * 24)), ]
    model <- lm(formula, data = train)

    # --------- TESTING ------------
    # getting the next 24 hours of data
    test <- data[(24 * d + 365 * 24 - 23):(24 * d + 365 * 24), ]
    predictions <- numeric(24)
    for (i in 1:24) {
        predictions[i] <- as.numeric(predict(model, newdata = test[i, ]))
        # Shift Y_d_1_24 to Y_predicted
        test[i + 1, ]$Y_d_1_24 <- predictions[i]
    }
    # Remove rows with NA values created by lagging
    test <- na.omit(test)

    test$y_hat <- predictions
    return(test)
}
```

```{r}
# Run the predictons for some days
predictions <- data.frame(matrix(ncol = 20, nrow = 0))
for (i in 1:21) {
    print(i)
    value <- predict_ar24(energy_load, i)
    predictions <- rbind(predictions, value)
}
```

```{r}
# Reshape the dataframe to long format
predictions_long <- pivot_longer(predictions, cols = c(load, y_hat), names_to = "type", values_to = "value")

# Plot
fig <- ggplot(predictions_long, aes(x = date, y = value, color = type)) +
    geom_line() + # Draw lines
    labs(x = "Date", y = "Value", title = "Load and Predicted Load Over Time - Expert Model") +
    theme_minimal() # Use a minimal theme for aesthetics
ggplotly(fig)

resids <- predictions$load - predictions$y_hat

hist(resids, main = "Histogram of Residuals", xlab = "Residuals Expert Model")
```

### Model 5: TBATS

```{r}
# Take approx 2 weeks for training
demand <- msts(energy_load[8424:8760, ]$load, seasonal.period = c(24, 7 * 24))

# ---- TBATS ----
fit <- tbats(demand)
checkresiduals(fit)
plot(forecast(fit, h = 24))

# predictions <- as.numeric(forecast(fit, h = 24)$mean)
```

### Model 6: NeuralProphet
*Already implemented and working well...*